---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      fig.path = "README_figs/README-")
library("DDMEutils")
library("ggplot2")
library("knitr")
```
 
# DDMEutils - Repository of useful functions for daily analysis and R programming

## 1 `insp_outlier()`

Created by Gabriel Motta for outlier detection in time series, based on `anomalize` package. S3 methods are defined for `ts` and `data.frame` objects.

### Unidimensional time series

Undated object `x`:

```{r, echo=FALSE}
x <- rpois(100, 10); x[c(10,30)] <- c(50,65) ; x[c(50)] <- -10
qplot(seq_along(x), x, geom = "line", xlab = "Index", ylab = "Values")
```

The function call results in a `data.table` object. If `x` has no explicit dates, a default date sequence is created.

```{r}
x_modif <- insp_outlier(as.ts(x))
x_modif[49:55]
```

```{r}
x_modif %>% 
  melt(id = "DATE", measure = patterns("SERIES")) %>% 
  dplyr::mutate(variable = factor(variable, labels = c("Original","Imputed"))) %>% {.[]} %>% 
  ggplot(aes(DATE,value,color = variable)) +
  geom_line() +
  labs(x = "Index", y = "Values", color = "") + 
  theme(legend.position = "bottom")
```



### Multiple time series

When working with multiple time series, one has to pay attention to two additional arguments: `margin` and `idcol, dtcol`. 

* If the series are arranged by rows, then set `margin = 1` and `idcol = <integer specifying index column>`. If there is no index column, then set `idcol = 0` or `idcol = NULL` (default);
* If framed by columns, then set `margin = 2` and `dtcol = <integer specifying a date column>`. Likewise, `dtcol = 0` or `dtcol = NULL` if there is not an explicit date column. 

#### Time series framed in rows

Example:

```{r, echo = FALSE}
d1 = cbind(x,x + rnorm(100)) %>% t() %>% as.data.table() %>% .[, ID := 1:2] %>% setcolorder('ID') %>% {.[]}
d2 = data.table(DATE = seq(as.Date('2018-01-01'), by = 1, length.out = 100), SERIES1 = x, SERIES2 = round(rcauchy(100),4))
```


```{r}
d1[,1:10]

# 'Wide' output format
insp_outlier(d1, margin = 1, idcol = 1) %>% 
  dplyr::sample_n(10) %>% 
  {.[order(ID, PERIOD)]}

# 'Long' output format
insp_outlier(d1, margin = 1, out_format = "long") %>% 
  dplyr::sample_n(10) %>% 
  {.[order(ID, PERIOD)]}
```


#### Time series framed in cols 

Example:

```{r}
d2[1:5]

# 'Wide' output format
insp_outlier(d2, margin = 2, dtcol = 1) %>% head()

# 'Long' output format
insp_outlier(d2, margin = 2, dtcol = 1, out_format = "long") %>% 
  dplyr::sample_n(10) %>% 
  {.[order(SERIES, DATE)]}
```


## 2 `insp_seasonality()`


## 3 Value replacements with `subs_any()` and `subs_na()`

### `subs_na()`

Replaces NA occurrences in variables of a table. The syntax is either on the form `subs_na(data, col_1 = "?", col_2 = 999)` or `subs_na(data, list(col_1 = "?", col_2 = 999))`, where '?' and 999 are examples of replacement values. 

### Example with `airquality` dataset

```{r}
data <- airquality %>% as.data.table()
head(data)

data %>% 
  subs_na(Ozone = 999, Solar.R = 999)

```

Regular expressions for equal value replacement and multiple column matching are supported: 

```{r}
# Replaces NA by 999 at columns that contain '.' 
data %>% subs_na("\\." = 999)
```


### `subs_any()`

Replaces any `x` value for any `y` within `data.frame` variables. The syntax is `subs_any(data, col_1 = list(<value>, <input>), col_2 = list(<value>, <input>))`.

Just like `subs_na()`, refer to multiple columns with regular expressions is supported.

```{r}
data %>% 
  subs_any("^M|^D" = list(5, 10000))
```


## 4 Cleaning text variables and column names with `clean()`

Default replacements are 

```{r}

```


## 5 Helpers for dealing with NA: `na_prop()`

### `na_prop()`

Check total and proportion of missing values by `data.frame` variables. 

```{r}
# Default print
data %>% na_prop()

# Restricted. Useful to retrieve variables that present more than min_prop of NA.   
data %>% na_prop(min_prop = 0.1)
```