---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE,
                      fig.path = "README_figs/README-")
library("devtools")
library("ggplot2")
library("knitr")
load_all()
```
 
# DDMEutils - Repository of useful functions for daily analysis and R programming

## 1 `insp_outlier()`

Written by Gabriel Motta for outlier detection in time series, based on `anomalize` package. S3 methods are defined for `ts` and `data.frame` objects.

### Unidimensional time series

Undated object `x`:

```{r, echo=FALSE}
x <- rpois(100, 10); x[c(10,30)] <- c(50,65) ; x[c(50)] <- -10
qplot(seq_along(x), x, geom = "line", xlab = "Index", ylab = "Values")
```

The function call results in a `data.table` object. If `x` has no explicit dates, a default date sequence is created.

```{r}
x_modif <- insp_outlier(as.ts(x))
x_modif[49:55]
```

```{r}
x_modif %>% 
  melt(id = "DATE", measure = patterns("SERIES")) %>% 
  dplyr::mutate(variable = factor(variable, labels = c("Original","Imputed"))) %>% {.[]} %>% 
  ggplot(aes(DATE,value,color = variable)) +
  geom_line() +
  labs(x = "Index", y = "Values", color = "") + 
  theme(legend.position = "bottom")
```



### Multiple time series

When working with multiple time series, one has to pay attention to two additional arguments: `margin` and `idcol, dtcol`. 

* If the series are arranged by rows, then set `margin = 1` and `idcol = <integer specifying index column>`. If there is no index column, then set `idcol = 0` or `idcol = NULL` (default);
* If framed by columns, then set `margin = 2` and `dtcol = <integer specifying a date column>`. Likewise, `dtcol = 0` or `dtcol = NULL` if there is not an explicit date column. 

#### Time series framed in rows

Example:

```{r, echo = FALSE}
d1 = cbind(x,x + rnorm(100)) %>% t() %>% as.data.table() %>% .[, ID := 1:2] %>% setcolorder('ID') %>% {.[]}
d2 = data.table(DATE = seq(as.Date('2018-01-01'), by = 1, length.out = 100), SERIES1 = x, SERIES2 = round(rcauchy(100),4))
```


```{r}
d1[,1:10]

# 'Wide' output format
insp_outlier(d1, margin = 1, idcol = 1) %>% 
  dplyr::sample_n(10) %>% 
  {.[order(ID, PERIOD)]}

# 'Long' output format
insp_outlier(d1, margin = 1, out_format = "long") %>% 
  dplyr::sample_n(10) %>% 
  {.[order(ID, PERIOD)]}
```


#### Time series framed in cols 

Example:

```{r}
d2[1:5]

# 'Wide' output format
insp_outlier(d2, margin = 2, dtcol = 1) %>% head()

# 'Long' output format
insp_outlier(d2, margin = 2, dtcol = 1, out_format = "long") %>% 
  dplyr::sample_n(10) %>% 
  {.[order(SERIES, DATE)]}
```


## 2 `insp_seasonality()`

Written by Gabriel Motta for getting seasonality of time series, based on `forecast` package. S3 methods are defined for `ts` and `data.frame` objects.

Options for `data.frame`: (check `args(insp_outlier.data.frame)` for defaults)

* `trend` T/F - if time series have trend;
* `margin` 
    + `margin = 1` for series framed row-wise. In this case, specify as well an integer to an index column on `idcol`;
    + `margin = 2` for series framed col-wise. Specify an integer to an date column on `dtcol`;
* `periodicity` of the series 


### Example for `airquality` dataset:

```{r}
data <- airquality %>% as.data.table()
head(data)
setorder(data, Month, Day)

data[, .(Wind, Temp)] %>% 
  insp_seasonality(margin = 2, dtcol = 0, periodicity = 'day')
```

Considering we have daily observations and a total length no longer than a year for 'Wind' and 'Temp' variables, seasonality 365 means that no periods of seasonality were found.


## 3 Value replacements with `subs_any()` and `subs_na()`

### `subs_na()`

Replaces NA occurrences in variables of a table. The syntax is either on the form `subs_na(data, col_1 = "?", col_2 = 999)` or `subs_na(data, list(col_1 = "?", col_2 = 999))`, where '?' and 999 are examples of replacement values. 

### Example with `airquality` dataset

```{r}
data <- airquality %>% as.data.table()
head(data)

data %>% 
  subs_na(Ozone = 999, Solar.R = 999)

```

Regular expressions for equal value replacement and multiple column matching are supported: 

```{r}
# Replaces NA by 999 at columns that contain '.' 
data %>% subs_na("\\." = 999)
```


### `subs_any()`

Replaces any `x` value for any `y` within `data.frame` variables. The syntax is `subs_any(data, col_1 = list(<value>, <input>), col_2 = list(<value>, <input>))`.

Just like `subs_na()`, refer to multiple columns with regular expressions is supported.

```{r}
data %>% 
  subs_any("^M|^D" = list(5, 10000))
```


## 4 Cleaning text variables and column names with `clean()`

Methods for cleaning headers and text variables within data.frame. For `factor` objects, cleans levels attribute. Default replacements are stored in exported list `.dict`, where names `rm` and `undln` stand for 'remove' and 'underline' actions.

```{r}
.dict

to_clean <- c('Header with Nº of meas./hour(in measurement unit)', 
              '\u00c0rtificial p\u00fcnctuat\u00ead attribute')
to_clean

# S3 method for char 
clean(to_clean)
```

Additional arguments: 

* `keep` - vector of chars to remain unchanged;
* `add_repl` - named vector of additional replacements desired.

```{r}
clean(to_clean, keep ='º')

clean(to_clean, add_repl = c('H' = 'HHH'))
```

For `data.frame`, further logical options are:

* `col_names` - T/F for changes in table header;
* `vars` - T/F for changes in table variables;
* `byref` - T/F if replacements are made by reference in the table.

```{r}
df_to_clean <- data.table(
  '"quoted name"' = 1,
  'text' = LETTERS[1:5] %>% stringr::str_replace_all( c('C'='Ç', 'A' = '"An extract from a book"'))
)

df_to_clean

# default: col_names = T, vars = F
df_to_clean %>% 
  clean(add_repl = c('"' = '')) %>% 
  clean(vars = T, col_names = F, keep = '\\s')
```


## 5 Helpers for dealing with NA: `na_prop()` and `na_input()`

### `na_prop()`

Check total and proportion of missing values by `data.frame` variables. 

```{r}
# Default print
data %>% na_prop()

# Restricted. Useful to retrieve variables that present more than min_prop of NA.   
data %>% na_prop(min_prop = 0.1)
```

### `na_input()`

Numerical imputation of vectors. Supported types:

* Aggregating functions: adjustable-window mean or median;
* Linear interpolation;
* Cubic spline interpolation;
* LOCF (Last Observation Carried Forward);
* NOCB (Next Observation Carried Backward).

```{r}
x <- c(NA, 10, 9, 8, 7, NA, NA, 100, 50)

# mean or median, controling window size
na_input(x, how = 'mean') # window default = Inf
na_input(x, how = 'median', window = 3)

# locf, nocb
na_input(x, how = 'locf')
na_input(x, how = 'nocb')

# Approximations
na_input(x, how = 'lin_interp')
na_input(x, how = 'cub_spline')
```

